Reasoning About a Machine with Local Capabilities

### 2018-05-11 toplas/paper.tex
Very first draft for awkward example proof sketch

\begin{lemma}
  \label{lem:correctness-g1}
  Let
\[
    \begin{array}{r@{}lr@{}l}
    c_{\var{adv}} & \;\defeq ((\rwx,\glob),\dots) & c_{g1} & \;\defeq ((\entry,\glob),\dots)
    \end{array}
\]
  and otherwise make assumptions about capabilities and linking similar to Lemma~\ref{lem:correctness-f1}.
  Then if $
  (\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step^* (\halted,m'),$
  then $m'(\flag) = 0$.
\end{lemma}

As explained in \sectionname~\ref{sec:stack-and-return-pointer}, the
macro-instruction \texttt{\footnotesize{reqglob $r_1$}} checks that the callback
is global, essentially to make sure it is not allocated on the stack where it
might contain old stack pointers or return pointers. Otherwise, the
encapsulation of our local stack frame could be broken. In the proof of
Lemma~\ref{lem:correctness-g1}, this requirement shows up because we invoke the
callback in a world that is only a private future world of the one where we
received the callback, precisely because we have invalidated the adversary's
local state (particularly their old stack and return capabilities). The callback
is still valid in this private future world, but only because we know that it is
global.

>>>>

\begin{lemma}
  \label{lem:correctness-g1}
  Let
\[
    \begin{array}{r@{}lr@{}l}
    c_{\var{adv}} & \;\defeq ((\rwx,\glob),\dots) & c_{g1} & \;\defeq ((\entry,\glob),\dots)
    \end{array}
\]
  and otherwise make assumptions about capabilities and linking similar to Lemma~\ref{lem:correctness-f1}.
  Then if $
  (\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step^* (\halted,m'),$
  then $m'(\flag) = 0$.
\end{lemma}
\begin{toplas}
  \begin{proof}[Proof sketch]
    Define a world $W_1$ with the following regions
    \begin{itemize}
    \item A malloc region, $\iota_{\malloc,0}$.
    \item A permanent region for the linking table that only accepts $\ms_\link$ and requires everything to be in $\stdvr$.
    \item A stack region, $\iota^{\pwl}_{b_\stk,e_\stk}$.
    \item A adversary region, $\iota^{\nwl,p}_{b_\adv,e_\stk}$
    \item A permanent static region for the flag and \texttt{g1}, i.e.\ a region that only
      accepts $\ms_\flag$ and $\ms_{g1}$.
    \end{itemize}
    If we can show
    \begin{equation}
      (\reg_0[\pcreg \mapsto c_\adv,r_\stk\mapsto c_\stk,r_1\mapsto c_{g1}],\ms_\malloc \uplus \ms_\link \uplus \ms_\stk \uplus \ms_\adv\uplus \ms_\flag  \uplus \ms_{g1}) \in \observations(W_1),\label{eq:g1:main-O-res}
  \end{equation}
  \footnote{We will ignore step indexes in this proof sketch.}then the $\observations$-relation ensures that a successfully halting configuration terminates in a memory that respects a private future world of $W$.
  In $W$ the flag region is permanent, so it guarantees that it will remain unchanged.
  As described in \sectionname~\ref{sec:reasoning}, we use the FTLR (Theorem~\ref{thm:ftlr}) to reason about unknown code pointed to by $c_\adv$.
  With the standard no-write-local region chosen for the adversary, it is easy to show that the $\readCond{}$ and $\writeCond{}$ holds for $c_\adv$ which gives us $c_\adv \in \stder(W)$ by the FTLR.
  In order to get \ref{eq:g1:main-O-res}, we need to show that (a) the memory satisfies the world and (b) the register file is in the $\stdrr(W)$.
  We have defined the world, so there (almost) is a one-to-one correspondence between region and memory segment in the initial configuration, so (a) is easily shown.
  In order to show (b), we need to show $c_\stk \in \stdvr(W)$ and $c_{g1} \in \stdvr(W)$.
  the former follows easily from the choice of a permit-write-local region in $W$ for the stack.
  In order to argue the latter, we basically have to argue that $c_{g1}$, our closure generator, respects the world $W$.
  This amounts to showing that the closures generated by \texttt{g1} also respects the invariants.
  The reasoning concerning variable $x$ follows that of \citet{Dreyer:jfp12}, so we will ignore it in the remainder of the proof and focus on the reasoning related to this low-level scenario.
  The capability for the closure is a global enter capability that we call $c_{f4}$.
  The remainder of the proof amounts to showing that it is safe to return $c_{f4}$ to the adversary.
  The adversary can use $c_{f4}$ whenever, so all we may assume about the configuration that $c_{f4}$ is invoked in is that the register-file $\reg_1$ and memory $\ms_1$ satisfies a world $W_1$ that is a private future world of $W$ with regions prompted the closure generation.
  We have to show that the invocation respects $W_1$ which corresponds to showing $(\reg_1[\pcreg \mapsto \updatePcPerm{c_{f4}}],\ms_1) \in \observations(W_1)$.
  We know the first of the execution, so we use the anti-reduction lemma (Lemma~\ref{lem:anti-red-obs}) to reason about the execution up until the first call.
  During the first bit of execution, it is checked that the stack pointer has permission read/write-local/execute and that the callback is global.
  Based on the stack pointers permission, we know that the capability is local which means that the region that governs it must be temporary.
  This allows us to construct $W_2 \futurestr W_1$ where the stack region has been revoked (notice that it is also important for the reasoning about $x$ that this is a private future world).
  In $W_2$, we replace the stack region with a temporary static region that governs our stack frame including the private registers we push to the stack and the stack recovery code as well as a standard region $\iote^{pwl}$-region for the remainder of the stack.
  As the callback is global, it is safe to invoke it in a private future world of $W_1$ (even though the code is unknown, we do not need to use the FTLR because we may assume that the arguments for the invocation of $c_{f4}$ are safe).
  Notice that had the callback been local, this would not be the case.
  We also need to reason about the return pointer we provide in the invocation.
  The return pointer is a local capability, so we may assume that it is invoked in some $W_3 \futurewk W_2$.
  This means that none of the temporary regions have been revoked, so the regions we replaced the old stack region with are still there in $W_3$.
  For the sake of presentation, we went through the above reasoning, but we actually us Lemma~\ref{lem:scall-works} to reason about \texttt{scall}.
  To reason about the next bit of execution we know, we once again use Lemma~\ref{lem:anti-red-obs} followed by \ref{lem:scall-works} to handle \texttt{scall}.
  As the private stack has changed, we need to replace the temporary static region that governs it, so it is revoked.
  This means that the \texttt{scall} takes place in a world $W_4 \futurestr W_3$.
  The reasoning about \texttt{scall} is the same as for the first callback invocation, so again we will get a world $W_5 \futurewk W_4$ in which we need to argue that the remainder of the execution is safe.
  At this point, we can use Lemma~\ref{lem:anti-red-obs} to get to the point where we jump to the return pointer.
  If we can argue that it is safe to jump to the return pointer we got initially from the adversary, then the proof is done.
  We have made no checks on the return pointer, so we have no idea whether it is local or global\footnote{We may assume that it is a capability that is executable when jumped to otherwise the execution fails which is considered safe.}.
  W.l.o.g.\ assume the return pointer is local.
  This means that we have to use it in a world $W_6$ that is a public future world of $W_1$, which means that we have to restore all the regions in the world that the adversary relies on.
  In particular, we will have to restore the stack region from $W_1$.
  We cannot just pick $W_1$ because Lemma~\ref{lem:anti-red-obs} requires $W_6 \futurestr W_5$.
  Before we proceed, let us consider where each of the worlds came from so far: As illustrated in \figurename~\ref{fig:worlds-in-pf-g1}, we chose $W_2$ and $W_4$ which means that we know exactly what changes they made.
  In particular, we did not mask any of the revoked or temporary regions in $W_1$ with a permanent region.
  This means that in $W_6$, we can reinstate every temporary and revoked region in $W_1$.
  In the private future worlds we construct, we only take transitions that relative to $W_1$ are public.
  The worlds we are given, $W_3$ and $W_5$ may have taken arbitrary public transitions, but this is not problem with respect to the public future world relation.
  In $W_5$ there will be regions to handle the stack.
  These regions obviously need to be revoked as the $W_1$ stack regions replace them.
  The non-stack regions in $W_5$ that are not present in $W_1$ are also added in $W_6$, which the public future world permits as it is extensional.
  All in all, we create a world in which the invariants that the adversary relies on has been restored which makes it safe to invoke the return pointer.
  \end{proof}
\end{toplas}
As explained in \sectionname~\ref{sec:stack-and-return-pointer}, the
macro-instruction \texttt{\footnotesize{reqglob $r_1$}} checks that the callback
is global, essentially to make sure it is not allocated on the stack where it
might contain old stack pointers or return pointers. Otherwise, the
encapsulation of our local stack frame could be broken. In the proof of
Lemma~\ref{lem:correctness-g1}, this requirement shows up because we invoke the
callback in a world that is only a private future world of the one where we
received the callback, precisely because we have invalidated the adversary's
local state (particularly their old stack and return capabilities). The callback
is still valid in this private future world, but only because we know that it is
global.
